CYO Project Report - Brazilian Ecommerce

### 2021-07-13 finalReport.Rmd
fine tuned code for final model and predictions

@@

weight_to_use <- mean(weight_to_use_per_partition)  
# 3
```

@@

weight_to_use <- mean(weight_to_use_per_partition) 
# 20
```

@@

weight_to_use <- mean(weight_to_use_per_partition) 
# 5/3 = 1.66667
```

@@

```{r Glm model}
# Let's try with glm model
set.seed(seed=1,sample.kind = 'Rounding')
tuningControl=trainControl(method = "cv", number = 2, p = .8)
glm_fit <- train(number_of_hours~., train_subset, method='glm',trControl=tuningControl)

glm_fit

predictions <- predict(glm_fit, test_subset %>% select(-number_of_hours))

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 198.8575 (improves if we use seller_id as predictor as well)
```
Warning:
prediction from a rank-deficient fit may be misleading

```{r GamLoess model, echo = FALSE}
set.seed(seed=1,sample.kind = 'Rounding')
tuningControl=trainControl(method = "cv", number = 3, p = .8)
#grid <- expand.grid(span = seq(0.15, 0.65, len = 10), degree = 1)
grid <- expand.grid(span = seq(0.15, 0.65, len = 3), degree = 1)
gamLoess_fit <- train(number_of_hours~., train_subset, method='gamLoess',tuneGrid=grid,trControl=tuningControl)

gamLoess_fit

predictions <- predict(gamLoess_fit,test_subset %>% select(-number_of_hours))

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
#194.6215

save(gamLoess_fit,file='rda/gamLoess_model.rda')
```

Warnings received:
extrapolation not allowed with blending
 (Discovered by lowesd)liv too small.    
 (Discovered by lowesd)lv too small
 extrapolation not allowed with blending
 non-list contrasts argument ignored

```{r GamLoess model with scaling, echo = FALSE}
# TODO review-keep or remove it
set.seed(seed=1,sample.kind = 'Rounding')
scaled_train_subset <- train_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale)) 
scaled_test_subset <- test_subset %>% mutate(across(c(mean_delay_in_handover,distance_between_two_points,volume,product_weight_g,freight_value,price),scale))

tuningControl=trainControl(method = "cv", number = 3, p = .8)
#grid <- expand.grid(span = seq(0.15, 0.65, len = 10), degree = 1)
grid <- expand.grid(span = seq(0.15, 0.65, len = 3), degree = 1)
gamLoess_fit <- train(number_of_hours~., scaled_train_subset, method='gamLoess',tuneGrid=grid,trControl=tuningControl)

gamLoess_fit

predictions <- predict(gamLoess_fit,scaled_test_subset %>% select(-number_of_hours))

RMSE(predictions,(scaled_test_subset %>% pull(number_of_hours) ))
#194.7017

save(gamLoess_fit,file='rda/gamLoess_model.rda')
```
Warnings received:
extrapolation not allowed with blending
 (Discovered by lowesd)liv too small.    
 (Discovered by lowesd)lv too small
 extrapolation not allowed with blending
 non-list contrasts argument ignored

@@

penalty_data <- training_dataset %>% get_delay_in_handover_to_carrier_per_seller_id()

training_dataset <- training_dataset %>% left_join(penalty_data,by='seller_id') %>% select(-seller_id,-handed_over_to_carrier_date,-shipping_limit_date)

mtry_to_use <- rf_fit$finalModel$mtry
ntree_to_use <- 500 #rf_fit$finalModel$ntree
tuningControl <- trainControl(method = "cv", number = 2, p = .8)
final_model <- train(number_of_hours~., training_dataset, method='rf',tuneGrid = data.frame(mtry = c(mtry_to_use)),trControl=tuningControl,ntree=ntree_to_use)

# final_model <- randomForest(number_of_hours~., training_dataset,mtry= mtry_to_use, ntree=ntree_to_use)

@@

```{r Final movel validation, echo = FALSE}
validation_dataset <- validation_dataset %>% compute_volume_using_product_info() %>% extract_distance_between_seller_customer() %>% left_join(penalty_data,by='seller_id') %>% select(-seller_id,-handed_over_to_carrier_date,-shipping_limit_date)

@@

## Explore seller_zip_code_prefix with number_of_hours

```{r Explore seller_zip_code_prefix with number_of_hours, echo = FALSE}
# TODO anyway to add regularization or remove the low impact seller_zip_code_prefix from the model
# let's plot errorbar for number of hours vs seller_zip_code_prefix
sellers_zip_of_interest <- training_dataset %>% dplyr::count(seller_zip_code_prefix) %>% filter(n>10) %>% pull(seller_zip_code_prefix)

training_dataset %>% filter(seller_zip_code_prefix %in% sellers_zip_of_interest) %>% group_by(seller_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(seller_zip_code_prefix=reorder(seller_zip_code_prefix,mean)) %>%  ggplot(aes(x=seller_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt() 
```
Indicates role between seller_zip_code and number of hours for delivery

## Explore customer_zip_code_prefix with number_of_hours

```{r Explore customer_zip_code_prefix with number_of_hours, echo=FALSE}
# TODO anyway to add regularization or remove the low impact customer_zip_code_prefix from the model
# let's plot errorbar for number of hours vs customer_zip_code_prefix
customers_zip_of_interest <- training_dataset %>% dplyr::count(customer_zip_code_prefix) %>% filter(n>=5) %>% pull(customer_zip_code_prefix)

training_dataset %>% filter(customer_zip_code_prefix %in% customers_zip_of_interest) %>% group_by(customer_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(customer_zip_code_prefix=reorder(customer_zip_code_prefix,mean)) %>%  ggplot(aes(x=customer_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt()
```
customer_zip_code_prefix can  be seen to be  influence number of hours for delivery

## Explore product_id with number_of_hours

```{r Explore product_id with number_of_hours}
# TODO anyway to add regularization or remove the low impact products from the model
# let's plot errorbar for number of hours vs product_id
products_of_interest <- training_dataset %>% dplyr::count(product_id)  %>% filter(n>=10) %>% pull(product_id)

training_dataset %>% filter(product_id %in% products_of_interest) %>% group_by(product_id) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_id=reorder(product_id,mean)) %>%  ggplot(aes(x=product_id,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_blank()) 
```
It influences the number of hours

====TODO review above

### Linear model trial and analysis

```{r Linear model, eval=FALSE, echo = FALSE}
# trying to reduce coefficients.
# in dimensions we can use volume to decrase predictors

########### only price predictor#############
## only price complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(price,number_of_hours), method='lm')

lm_fit

predictions <- predict(lm_fit,test_subset %>% select(price))

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 220.4085

## only price,distance_between_two_points complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(price,distance_between_two_points,number_of_hours), method='lm')

lm_fit

predictions <- predict(lm_fit,test_subset %>% select(price,distance_between_two_points))

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 202.0175

## only price,distance_between_two_points,product_weight_g complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(price,distance_between_two_points,product_weight_g,number_of_hours), method='lm')

lm_fit

predictions <- predict(lm_fit,test_subset %>% select(price,distance_between_two_points,product_weight_g))

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 201.1412

## only price,distance_between_two_points,product_weight_g,volume complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(price,distance_between_two_points,product_weight_g,volume,number_of_hours), method='lm')

lm_fit

predictions <- predict(lm_fit,test_subset %>% select(price,distance_between_two_points,product_weight_g,volume))

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 201.1086

## only price,distance_between_two_points,product_weight_g,volume,freight_value complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(price,distance_between_two_points,product_weight_g,volume,freight_value,number_of_hours), method='lm')

lm_fit

predictions <- predict(lm_fit,test_subset %>% select(price,distance_between_two_points,product_weight_g,volume,freight_value))

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 201.1047

## only product_category_name complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(product_category_name,number_of_hours), method='lm')

lm_fit

predictions <- predict(lm_fit,test_subset %>% select(product_category_name))

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 218.8943

# if we combine this with other predictors
## only price,distance_between_two_points,product_category_name complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(price,distance_between_two_points,product_category_name,number_of_hours), method='lm')

lm_fit

predictions <- predict(lm_fit,test_subset %>% select(price,distance_between_two_points,product_category_name))

RMSE(predictions,(test_subset %>% pull(number_of_hours) ))
# 199.8844
```

```{r Linear model with all predictors, echo=FALSE}
# above is a fine model
# with all predictors
lm_fit <- train(number_of_hours~., train_subset, method='lm')

lm_fit

predictions <- predict(lm_fit,test_subset %>% select(-number_of_hours))

RMSE((test_subset %>% pull(number_of_hours) ), predictions)
# 199.4791

save(lm_fit,file='rda/lm_model.rda')
```
Warnings received:
prediction from a rank-deficient fit may be misleading

To improve model performance consider--> changing character categorical variables to numerical categorical variables like

  product_category_name=as.factor(as.numeric(product_category_name))

use levels from original data for completeness

@@

## Experimets which failed

Let's play with seller_id. It deriorates rf model and for gam model there was no improvement
```{r play with seller_id, eval=FALSE, echo = FALSE}
set.seed(1,sample.kind = 'Rounding')
test_indices <- createDataPartition(wrangled_data$number_of_hours,times=1,p=0.2,list=FALSE)
dummy_data <- wrangled_data[-test_indices,]
dummy_data <- dummy_data %>% mutate(volume=product_length_cm*product_height_cm*product_width_cm) %>% mutate(seller_id=factor(as.numeric(seller_id))) %>% select(-customer_unique_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix,-product_length_cm,-product_height_cm,-product_width_cm)

# used gam model but it did not use seller id at all 
# Let's try with rf model
set.seed(seed=1,sample.kind = 'Rounding')
tuningControl=trainControl(method = "cv", number = 2, p = .8)
rf_fit_dummy <- train(number_of_hours~., dummy_train_subset, method='rf',tuneGrid = data.frame(mtry = seq(5, 7, 1)),trControl=tuningControl) # nodesize = 50,maxnodes=2000,ntree=200

predictions <- predict(rf_fit_dummy,dummy_test_subset %>% select(-number_of_hours))

RMSE(predictions,(dummy_test_subset %>% pull(number_of_hours) ))
# 203.6908

# Let's try with glm model
set.seed(seed=1,sample.kind = 'Rounding')
tuningControl=trainControl(method = "cv", number = 2, p = .8)
glm_fit_dummy <- train(number_of_hours~., dummy_train_subset, method='glm',trControl=tuningControl)

glm_fit_dummy

predictions <- predict(glm_fit_dummy,dummy_test_subset %>% select(-number_of_hours))

RMSE(predictions,(dummy_test_subset %>% pull(number_of_hours) ))
# 195.4884
```

>>>>

```{r Dataframe to store target encoding weights, include=FALSE}
target_encoding_weights <- data.frame(effect=character(0),weight=double(0)) 
```

@@

weight_to_use <- mean(weight_to_use_per_partition)  
# 3

target_encoding_weights <- target_encoding_weights %>% add_row(effect='seller_id',weight=weight_to_use)
```

@@

weight_to_use <- mean(weight_to_use_per_partition) 
# 20

target_encoding_weights <- target_encoding_weights %>% add_row(effect='customer_state',weight=weight_to_use)
```

@@

weight_to_use <- mean(weight_to_use_per_partition) 
# 5/3 = 1.66667

target_encoding_weights <- target_encoding_weights %>% add_row(effect='customer_loc',weight=weight_to_use)
```

@@

```{r Prepare seller_id effect based on target encoding, include = FALSE}
# prepare seller_id_effect based on regularisation weight we have identified
prepare_seller_id_effect <- function(train_df, weight_to_use) {
    global_mean <- mean(train_df$number_of_hours)

    set.seed(1,sample.kind='Rounding')
    test_weight_indices <- createDataPartition(train_df$number_of_hours, p=0.2,times=10,list=FALSE)

    seller_id_effects_list <- lapply(1:10, function(index) {
      train_weight_set <- train_df[-test_weight_indices[,index],]

      seller_id_effects
    })

    sie <- seller_id_effects_list[[1]]
    for (index in 2:10) {
       sie <- sie %>% 
         full_join(seller_id_effects_list[[index]],by='seller_id', suffix=c("1",index))
    }

    #seller id effects are ready
    seller_id_effects
}
```

```{r Prepare customer_state effect based on target encoding, include = FALSE}
# prepare consumer_state_effect based on regularisation weight we have identified
prepare_customer_state_effect <- function(train_df, weight_to_use) {
  global_mean <- mean(train_df$number_of_hours)

  set.seed(1,sample.kind='Rounding')
  test_weight_indices <- createDataPartition(train_df$number_of_hours,p=0.2,times=10,list=FALSE)

  customer_state_effects_list <- lapply(1:10, function(index) {
      train_weight_set <- train_df[-test_weight_indices[,index],]

  cse <- customer_state_effects_list[[1]]
  for (index in 2:10) {
     cse <- cse %>% 
       full_join(customer_state_effects_list[[index]],by='customer_state', suffix=c("1",index))
  }

  #customer state effects are ready
  customer_state_effects
}
```

```{r Prepare customer_loc effect based on target encoding, include = FALSE}
# prepare customer_loc_effect based on regularisation weight we have identified
prepare_customer_loc_effect <- function(train_df, weight_to_use) {
  global_mean <- mean(train_df$number_of_hours)

  set.seed(1,sample.kind='Rounding')
  test_weight_indices <- createDataPartition(train_df$number_of_hours,p=0.2,times=10,list=FALSE)

  customer_loc_effects_list <- lapply(1:10, function(index) {
      train_weight_set <- train_df[-test_weight_indices[,index],]

  # where-ever value is NA, we have used global mean
  cle <- cle %>% mutate(across(-customer_loc,~if_else(is.na(.x),global_mean,.x)))

  #customer loc effects are ready
  customer_loc_effects
}
```

```{r Enhance training and test dataset for seller_id effect based on target encoding, include = FALSE}
enhance_datasets_using_seller_id_effect <- function(seller_id_effects,train_df,test_df,cleanup) {
  #we will use seller id effects computed using training data in test data.
  train_df <- train_df %>% left_join(seller_id_effects,by='seller_id') %>% mutate(seller_id_effect=if_else(is.na(seller_id_effect),global_mean,seller_id_effect))

  test_df <- test_df %>% left_join(seller_id_effects,by='seller_id')

  # cleanup
  if (cleanup) {
    train_df <- train_df %>% select(-product_category_name,-seller_id)
    test_df <- test_df %>% select(-product_category_name,-seller_id)
  }

  list(train_df,test_df)
}
```

```{r Enhance training and test dataset for customer_state effect based on target encoding, include = FALSE}
enhance_datasets_using_customer_state_effect <- function(customer_state_effects,train_df,test_df,cleanup) {
  #we will use customer state effects computed using training data in test data.
  train_df <- train_df %>% left_join(customer_state_effects,by='customer_state') %>%
    mutate(customer_state_effect=if_else(is.na(customer_state_effect),
                                         global_mean,
                                         customer_state_effect))

  test_df <- test_df %>% left_join(customer_state_effects,by='customer_state')

  # cleanup
  if (cleanup) {
    train_df <- train_df %>% select(-customer_state)
    test_df <- test_df %>% select(-customer_state)
  }

  list(train_df,test_df)
}
```

```{r Enhance training and test dataset for customer_loc effect based on target encoding, include = FALSE}
enhance_datasets_using_customer_loc_effect <- function(customer_loc_effects,train_df,test_df,cleanup) {
  #we will use customer loc effects computed using training data in test data.
  train_df <- train_df %>% 
    left_join(customer_loc_effects,by='customer_loc') %>%
    mutate(customer_loc_effect=if_else(is.na(customer_loc_effect),global_mean,customer_loc_effect))

  test_df <- test_df %>% left_join(customer_loc_effects,by='customer_loc')

  # cleanup
  if (cleanup) {
    train_df <- train_df %>% select(-customer_loc,-customer_zip_code_prefix)
    test_df <- test_df %>% select(-customer_loc,-customer_zip_code_prefix)
  }

  list(train_df,test_df)
}
```

```{r Prepare data set for training, include = FALSE}
# prepares data for training and testing
# returns list of data frames. First list has training dataframe. Second list has test data frame
prepare_data_for_training_testing <- function(train_ds, test_ds, cleanup) {
  print('Adding column for delay in handover')
  penalty_data <- train_ds %>% get_delay_in_handover_to_carrier_per_seller_id()

  #we will use penalty_data computed using training data in test data.
  train_ds <- train_ds %>% left_join(penalty_data,by='seller_id') 
  test_ds <- test_ds %>% left_join(penalty_data,by='seller_id')
  if (cleanup) {
    train_ds <- train_ds %>% select(-handed_over_to_carrier_date,-shipping_limit_date)
    test_ds <- test_ds %>% select(-handed_over_to_carrier_date,-shipping_limit_date)
  }

  print('Enhancing data for seller id effects')
  weight_to_use <- target_encoding_weights %>% filter(effect=='seller_id') %>% pull(weight)
  seller_id_effects <- prepare_seller_id_effect(train_ds,weight_to_use)
  dataset_list <- enhance_datasets_using_seller_id_effect(seller_id_effects,train_ds,test_ds,cleanup)
  train_ds <- dataset_list[[1]]
  test_ds <- dataset_list[[2]]

  print('Enhancing data for customer state effects')
  weight_to_use <- target_encoding_weights %>% filter(effect=='customer_state') %>% pull(weight)
  customer_state_effects <- prepare_customer_state_effect(train_ds,weight_to_use)
  dataset_list <- enhance_datasets_using_customer_state_effect(customer_state_effects,train_ds,test_ds,cleanup)
  train_ds <- dataset_list[[1]]
  test_ds <- dataset_list[[2]]

  print('Enhancing data for customer loc effects')
  weight_to_use <- target_encoding_weights %>% filter(effect=='customer_loc') %>% pull(weight)
  customer_loc_effects <- prepare_customer_loc_effect(train_ds,weight_to_use)
  dataset_list <- enhance_datasets_using_customer_loc_effect(customer_loc_effects,train_ds,test_ds,cleanup)
  dataset_list
}
```

```{r Prepare dataset for final model training, echo = FALSE}
set.seed(seed=1,sample.kind = 'Rounding')

validation_dataset <- validation_dataset %>% 
  compute_volume_using_product_info() %>% 
  extract_distance_between_seller_customer() %>% 
  compute_rough_customer_location()

dataset_list <- prepare_data_for_training_testing(training_dataset,validation_dataset,FALSE)
training_dataset <- dataset_list[[1]]
validation_dataset <- dataset_list[[2]]

validation_dataset <- validation_dataset %>%
  semi_join(training_dataset,by='product_category_name') %>% 
  semi_join(training_dataset,by='seller_id') %>% 
  semi_join(training_dataset,by='customer_state') %>%
  semi_join(training_dataset,by='customer_loc')

# cleanup
training_dataset <- training_dataset %>%
  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix)
validation_dataset <- validation_dataset %>%
  select(-handed_over_to_carrier_date,-shipping_limit_date,-product_category_name,-seller_id,-customer_state,-customer_loc,-customer_zip_code_prefix)
```

@@

mtry_to_use <- rf_fit$finalModel$mtry
ntree_to_use <- 500 #rf_fit$finalModel$ntree
#tuningControl <- trainControl(method = "cv", number = 2, p = .8)
#final_model <- train(number_of_hours~., training_dataset, method='rf',tuneGrid = data.frame(mtry = c(mtry_to_use)),trControl=tuningControl,ntree=ntree_to_use)

final_model <- randomForest(number_of_hours~., training_dataset,mtry= mtry_to_use, ntree=ntree_to_use)

@@

```{r Final movel validation, echo = FALSE}
validation_dataset <- validation_dataset %>% select(-seller_id,-handed_over_to_carrier_date,-shipping_limit_date)

@@

===TODO review

@@
